Re-examining cross-cultural similarity judgements using lexical co-occurrence

### 2023-01-19 code/4_paper/manuscript/manuscript.Rmd
small changes based on Mike's feedback

abstract: >
 Is “cow” more closely related to “grass” or “chicken”? Speakers of different languages judge similarity in this context differently, but why? One possibility is that cultures co-varying with these languages induce variation in conceptualizations of similarity. Specifically, East Asian cultures may promote reasoning about thematic similarity, by which cow and grass are more related whereas Western cultures may bias similarity judgements toward taxonomic relations, like cow-chicken. This difference in notions of similarity is the consensus interpretation for cross-cultural variation in this paradigm. We consider, and provide evidence for, an alternative possibility, by which notions of similarity are equivalent across contexts, but the statistics of the environment vary. On this account, similarity judgements are guided by co-occurrence in experience, and observing or hearing about cows and grass or cows and chickens more often could induce preferences for the relevant grouping, and account for apparent differences in notions of similarity across contexts.

@@

## Language as a proxy of variation in experience
While co-occurrences in experience are difficult to measure, co-occurrence in language can provide a rough proxy--and indeed, language may afford many of the “experiences” that people have with infrequently encountered items, like cows or helicopters. In this study, we use co-occurrence in language as a proxy of cultural experience. This is a generalizing assumption, and while no linguistic corpus can be expected to fully capture a culture, using this proxy as a model of culture provides a conservative test of the hypothesis that differences in the input, rather than the process, of reasoning produces cultural variation in this task. This means that if such a model does work -- if linguistic co-occurrence can predict cultural variation in this similarity task without building differences in the decision process into the model -- it is relatively strong evidence that language alone (perhaps as a more easily observed source of information on broader cultural and ecological variation) may account for these cross-cultural differences in reasoning.[^footnote]

@@

<!-- Additionally, outside of language, cross-cultural differences could come from a couple different mechanisms:  -->
@Ji2004 suggest that the notion of similarity is what varies across cultural groups, but it is quite possible that similarity is regarded in the same way and it is actually the input to this similarity judgment that varies across culture contexts. It might be possible to gain traction on possible mechanisms by examining whether variation in similarity judgments co-varies with environmental statistics that differ across the cultural and linguistic contexts.

In this study, we ask whether variation in lexical co-occurrence can account for the cross-context (culture and language) differences observed in how people evaluate similarity. First, we attempt to replicate @Ji2004 by running a study comparing English speakers in the US and Mandarin speakers in mainland China. Second, we evaluate whether the differences observed between English and Mandarin speakers’ similarity judgments also extend to comparisons between English and Vietnamese, by running a similar study between English speakers in the US and Vietnamese speakers in Vietnam. Vietnam is a Southeast Asian country that borders China and is historically greatly influenced by Chinese culture [@Hui2002]. Therefore, it serves as a suitable cultural context to investigate whether the claim made by @Ji2004 and previous studies, that Eastern and Western cultures have different notions of similarity, extends beyond mainland China.

@@

# Methods 
```{r get responding and demographic data, include = F}
df <- read.csv("../../../data/data_USCNVN_ENZHVI.csv")

@@

triads_omit <- read.csv("../../../data/triads_omit.csv") %>%
  pull(x)
```

@@

We excluded `r df.excl_US$ppts_finished - df.excl_US$ppts_after_att_check_excl` US participants, `r df.excl_VN$ppts_finished - df.excl_VN$ppts_after_att_check_excl` Vietnam participants, and `r df.excl_CN$ppts_finished - df.excl_CN$ppts_after_att_check_excl` China participants who did not answer all attention checks correctly. We followed 4 exclusion criteria that aim to retain only participants who are influenced by one culture: (1) non-native speakers of English and Vietnamese, respectively, (2) fluent in at least one of the other two study languages (Vietnamese for US participants, English for Vietnamese participants and Chinese participants), (3) have lived outside of the test country (US, Vietnam, or China) for more than two years, and (4) have significant international experience (more than 6 international experiences of 2 days or longer.) We did not use a particular criterion for a language if it would exclude 25% or more of any one sample. In this round of exclusion, we excluded `r df.excl_US$ppts_after_att_check_excl - df.excl_US$ppts_after_demog_excl` US, `r df.excl_VN$ppts_after_att_check_excl - df.excl_VN$ppts_after_demog_excl` Vietnam participants, and `r df.excl_CN$ppts_after_att_check_excl - df.excl_CN$ppts_after_demog_excl` China participants. After these exclusions, the US sample included `r df.excl_US$ppts_after_demog_excl` participants (`r df.gender_E %>% filter(gender == "Male") %>% pull(n)`M, `r df.gender_E %>% filter(gender == "Female") %>% pull(n)`F, `r df.gender_E %>% filter(gender == "Non-binary") %>% pull(n)` non-binary, `r df.gender_E %>% filter(gender == "Decline to answer") %>% pull(n)` other), with mean age = `r round(df.age_E$mean_age, 2)` (SD = `r round(df.age_E$sd_age, 2)`) and median age = `r df.age_E$median_age`. The Vietnam sample included `r df.excl_VN$ppts_after_demog_excl` participants (`r df.gender_V %>% filter(gender == "Nam") %>% pull(n)`M, `r df.gender_V %>% filter(gender == "Nữ") %>% pull(n)`F, `r df.gender_V %>% filter(gender == "Từ chối trả lời") %>% pull(n)` other), with mean age = `r round(df.age_V$mean_age, 2)` (SD = `r round(df.age_V$sd_age, 2)`) and median age = `r df.age_V$median_age`. The China sample included `r df.excl_CN$ppts_after_demog_excl` participants (`r df.gender_M %>% filter(gender == "男性") %>% pull(n)`M, `r df.gender_M %>% filter(gender == "女性") %>% pull(n)`F, `r df.gender_M %>% filter(gender == "拒绝回答") %>% pull(n)` other), with mean age = `r round(df.age_M$mean_age, 2)` (SD = `r round(df.age_M$sd_age, 2)`) and median age = `r df.age_M$median_age`.

Notably, we lost a majority of our Vietnam participants (more than 60%) in the attention check exclusion. One reason we suspect why this might have happened is because our Vietnam participants are less familiar with research surveys and attention check questions, and thus might have thought too much about the attention check questions. We carried out an exploratory analysis where we used a less stringent attention check exclusion (covered in the final part of the Results & Analysis section). To preview our findings for this analysis, we did not find any difference in significant results when compared with the main analysis.

@@

TODO; ADD INFO ABOUT MANDARIN CORPUS

@@

fit.country_EN_VN = glmer(responses_theme ~ country + (1 | subject) + (country | triad), 
                    data = df %>% filter(country != "China"),
                    family = "binomial")
summary(fit.country_EN_VN) 
```
The group means of proportion of thematic response in mainland China is the highest (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by the groups means in Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), which is slightly higher than that of the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`).

```{r echo=FALSE, warning=FALSE, cache = TRUE, fig.cap="Proportion of thematic responses by country. Only US-China responding comparison shows a siginficant difference. We could not extend this to US-Vietnam responding comparison."}

#show violin plot
ggplot(df.country,
       mapping = aes(x = factor(country, levels=c("China", "US", "Vietnam")), 
                     y = theme_resp_percent, 
                     color = factor(country, levels=c("China", "US", "Vietnam")))) +
  geom_violin() +
  geom_jitter(height = 0, 
              alpha = 0.3) +  
  stat_summary(fun.data = "mean_cl_boot", 
               geom = "pointrange") + 
  labs(y = "Proportion Thematic Chosen", 
       x = "Country", 
       color = "Country") + 
  scale_color_manual(values=c("#D63230", "#1C77C3", "#F39237"))
```

```{r correlation of country response, cache = TRUE, echo=FALSE, warning=FALSE}
#US-China, US-Vietnam, China-Vietnam
df.triad_country <- df %>%
  group_by(triad, country) %>%
  summarize(emp_theme_prop = mean(responses_theme))

df.triad_country <- pivot_wider(data = df.triad_country, 
                                names_from = country, 
                                values_from = emp_theme_prop)
plotUSCN = ggplot(data = df.triad_country,
                  mapping = aes(x = US, 
                                y = China, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1) + 
  theme(axis.title = element_text(size=rel(0.75)))

plotUSVN = ggplot(data = df.triad_country,
                  mapping = aes(x = US, 
                                y = Vietnam, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Chosen (Vietnam)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

plotCNVN = ggplot(data = df.triad_country,
                  mapping = aes(x = Vietnam,
                                y = China, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (Vietnam)", 
        y = "Proportion Thematic Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

corr.USCN = apa_print(cor.test(df.triad_country$US, 
                  df.triad_country$China, na.rm=T))

corr.USVN = apa_print(cor.test(df.triad_country$US, 
                  df.triad_country$Vietnam, na.rm=T))

corr.CNVN = apa_print(cor.test(df.triad_country$China, 
                  df.triad_country$Vietnam, na.rm=T))
```

<!-- With a Pearson correlation test, we observed high correlations between thematic responding across all three countries. The proportion of US thematic responding is strongly correlated with thematic responding in China, `r corr.USCN$full_result`. The proportion of US thematic responding is strongly correlated with thematic responding in Vietnam, `r corr.USVN$full_result`. Likewise, the proportion of China thematic response is strongly correlated with thematic responding in Vietnam, `r corr.CNVN$full_result`. This points to a strong signal of universality across cultural contexts regarding similarity judgments. -->

```{r USCN effect size comparison, cache = TRUE, include=FALSE}
USCN_cohen <- cohen.d(df.triad_country$China, df.triad_country$US)
```

@@

## 2. Can the differences in similarity judgments between English, Mandarin and Vietnamese speakers be explained by variation in lexical statistics?
### Is responding in each cultural context predicted by the corresponding lexical statistics (raw co-occurrences)? 
#### Raw lexical co-occurrences
```{r Pearson correlation between raw co-occurrences of corpora and responding, warning=FALSE, eval = F, cache = TRUE, include=FALSE}

#raw co-occurrences
df.triad_country_freq <- df %>%
  filter(!(triad %in% triads_omit)) %>% # omit triads that do not have meaningful prediction (at least 1 non-zero prediction in both languages)
  group_by(triad, country, theme_freq_prop_E, theme_freq_prop_V, theme_freq_prop_M) %>%  # not working 
  summarize(emp_theme_prop = mean(responses_theme))
df.triad_country_freq   #sanity

df.triad_country_freq_E <- df.triad_country_freq %>%
  filter(country == "US")

#plotting English corpus against US responding data
plotE = ggplot(data = df.triad_country_freq_E,
        mapping = aes(x = emp_theme_prop,
                     y = theme_freq_prop_E,
                     label = triad)) +
   geom_text() +
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)",
        y = "Proportion Thematic Predicted (English)")
plotE

df.triad_country_freq_V <- df.triad_country_freq %>%
  filter(country == "Vietnam")

#plotting Vietnamese corpus against VN responding data
plotV = ggplot(data = df.triad_country_freq_V,
        mapping = aes(x = emp_theme_prop,
                     y = theme_freq_prop_V,
                     label = triad)) +
   geom_text() +
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (Vietnam)",
        y = "Proportion Thematic Predicted (Vietnamese)")
plotV

df.triad_country_freq_M <- df.triad_country_freq %>%
  filter(country == "China")

#plotting China corpus against CN responding data
plotM = ggplot(data = df.triad_country_freq_M,
        mapping = aes(x = emp_theme_prop,
                     y = theme_freq_prop_M,
                     label = triad)) +
   geom_text() +
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (China)",
        y = "Proportion Thematic Predicted (Chinese)")
plotM

#Pearson correlation between corpus and responding
corr.E = apa_print(cor.test(df.triad_country_freq_E$emp_theme_prop,
                  df.triad_country_freq_E$theme_freq_prop_E, na.rm=T))
corr.E

corr.V = apa_print(cor.test(df.triad_country_freq_V$emp_theme_prop,
                  df.triad_country_freq_V$theme_freq_prop_V, na.rm=T))
corr.V

corr.M = apa_print(cor.test(df.triad_country_freq_M$emp_theme_prop,
                  df.triad_country_freq_M$theme_freq_prop_M, na.rm=T))
corr.M

#Pearson correlation between EN and VI corpus
corr.EV = apa_print(cor.test(df.triad_country_freq_V$theme_freq_prop_V,
                   df.triad_country_freq_E$theme_freq_prop_E, na.rm=T))
corr.EM = apa_print(cor.test(df.triad_country_freq_M$theme_freq_prop_M,
                   df.triad_country_freq_E$theme_freq_prop_E, na.rm=T))
corr.MV = apa_print(cor.test(df.triad_country_freq_V$theme_freq_prop_V,
                   df.triad_country_freq_M$theme_freq_prop_M, na.rm=T))
```

```{r echo=FALSE, warning = F, eval = F, cache = TRUE, paged.print=FALSE, fig.cap ="Correlation between English corpus lexical co-occurrence and US responding, and Vietnamese corpus lexical co-occurrence and Vietnam responding."}
plotE + plotV + plotM
  plot_layout(ncol = 3)
```

@@

We found that all corpora are significant predictors of all cultural context responding.
```{r US responding predicted by each corpus model raw co-occurrences, warning=FALSE, include=F, cache = TRUE}
# English 
fit.EN_US = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)

@@

#Mandarin
fit.ZH_US = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_US)
fit.ZH_US.anova = Anova(fit.ZH_US, type = 3)
```
For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH) corpus are all significant predictors. EN corpus: $\beta$ = `r round(fixef(fit.EN_US)["theme_freq_prop_E"], 2)`, $\chi^2$(`r fit.EN_US.anova["theme_freq_prop_E", "Df"]`) = `r round(fit.EN_US.anova["theme_freq_prop_E", "Chisq"], 2)`, p `r ifelse( fit.EN_US.anova["theme_freq_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.EN_US.anova["theme_freq_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.VI_US)["theme_freq_prop_V"], 2)`, $\chi^2$(`r fit.VI_US.anova["theme_freq_prop_V", "Df"]`) = `r round(fit.VI_US.anova["theme_freq_prop_V", "Chisq"], 2)`, p `r ifelse( fit.VI_US.anova["theme_freq_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.VI_US.anova["theme_freq_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ZH_US)["theme_freq_prop_M"], 2)`, $\chi^2$(`r fit.ZH_US.anova["theme_freq_prop_M", "Df"]`) = `r round(fit.ZH_US.anova["theme_freq_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ZH_US.anova["theme_freq_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ZH_US.anova["theme_freq_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

@@

#Mandarin
fit.ZH_VN = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_VN)
fit.ZH_VN.anova = Anova(fit.ZH_VN, type = 3)
```
For Vietnam responding: English (EN), Vietnamese (VI), and Mandarin (ZH) corpus are all significant predictors. EN corpus: $\beta$ = `r round(fixef(fit.EN_VN)["theme_freq_prop_E"], 2)`, $\chi^2$(`r fit.EN_VN.anova["theme_freq_prop_E", "Df"]`) = `r round(fit.EN_VN.anova["theme_freq_prop_E", "Chisq"], 2)`, p `r ifelse( fit.EN_VN.anova["theme_freq_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.EN_VN.anova["theme_freq_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.VI_VN)["theme_freq_prop_V"], 2)`, $\chi^2$(`r fit.VI_VN.anova["theme_freq_prop_V", "Df"]`) = `r round(fit.VI_VN.anova["theme_freq_prop_V", "Chisq"], 2)`, p `r ifelse( fit.VI_VN.anova["theme_freq_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.VI_VN.anova["theme_freq_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ZH_VN)["theme_freq_prop_M"], 2)`, $\chi^2$(`r fit.ZH_VN.anova["theme_freq_prop_M", "Df"]`) = `r round(fit.ZH_VN.anova["theme_freq_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ZH_VN.anova["theme_freq_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ZH_VN.anova["theme_freq_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

@@

@@

#### Cosine distances of fastText word vectors
Similar to the results from the analysis with raw lexical co-occurrences, we found that all corpora are significant predictors of all cultural context responding.

@@

#Mandarin
fit.ZH_US_cos = glmer(responses_theme ~ theme_cosine_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US"), 
          family="binomial")
summary(fit.ZH_US_cos)
fit.ZH_US_cos.anova = Anova(fit.ZH_US_cos, type = 3)
fit.ZH_US_cos.anova
```
For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH) corpus are significant predictors of US responding (EN-US model: $\beta$ = `r round(fixef(fit.EN_US_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.EN_US_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.EN_US_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.EN_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.EN_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`; VI-US model: $\beta$ = `r round(fixef(fit.VI_US_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.VI_US_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.VI_US_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p = `r ifelse( fit.VI_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.VI_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`; ZH-US model: $\beta$ = `r round(fixef(fit.ZH_US_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ZH_US_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ZH_US_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`).

@@

For Vietnam (VN) responding: English (EN), Vietnamese (VI), and Mandarin (ZH) corpus are significant predictors of VN responding (EN-VN model: $\beta$ = `r round(fixef(fit.EN_VN_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.EN_VN_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.EN_VN_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.EN_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.EN_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`; VI-VN model: $\beta$ = `r round(fixef(fit.VI_VN_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.VI_VN_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.VI_VN_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p = `r ifelse( fit.VI_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.VI_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`; ZH-VN model: $\beta$ = `r round(fixef(fit.ZH_VN_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ZH_VN_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ZH_VN_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`).

@@

#### Raw co-occurrences 
When we included all corpora thematic cosine distance proportion as predictors, all corpora are significant in predicting all cultural context's responding.

@@

We observed some level of language specificity from this analysis. The English corpus is the best predictor for US responding, and the Mandarin corpus is the best predictor for China response. While this is not the case with the Vietnamese corpus and the Vietnam responding, the Vietnamese corpus is still a significant predictor for the Vietnam responding.

>>>>

abstract: >
 Is “cow” more closely related to “grass” or “chicken”? Speakers of different languages judge similarity in this context differently, but why? One possibility is that cultures co-varying with these languages induce variation in conceptualizations of similarity. Specifically, East Asian cultures may promote reasoning about thematic similarity, by which cow and grass are more related, whereas Western cultures may bias similarity judgements toward taxonomic relations, like cow-chicken. This difference in notions of similarity is the consensus interpretation for cross-cultural variation in this paradigm. We consider, and provide evidence for, an alternative possibility, by which notions of similarity are equivalent across contexts, but the statistics of the environment vary. On this account, similarity judgements are guided by co-occurrence in experience, and observing or hearing about cows and grass or cows and chickens more often could induce preferences for the relevant grouping, and account for apparent differences in notions of similarity across contexts.

@@

## Language as a proxy of variation in experience
While co-occurrences in experience are difficult to measure, co-occurrence in language can provide a rough proxy--and indeed, language may afford many of the “experiences” that people have with infrequently encountered items, like cows or helicopters. In this study, we use co-occurrence in language as a proxy of cultural experience. This is a generalizing assumption, and while no linguistic corpus can be expected to fully capture a culture, using this proxy as a model of culture provides a conservative test of the hypothesis that differences in the input, rather than the process, of reasoning produces cultural variation in this task. This means that if such a model does work -- if linguistic co-occurrence can predict cultural variation in this similarity task without building differences in the decision process into the model -- it is relatively strong evidence that language alone (perhaps as a more easily observed source of information on broader cultural and ecological variation) may account for these cross-cultural differences in reasoning.

While we discuss cross-cultural variability at the level of countries or larger world areas, these are not cultural monoliths. For convenience, we operationalize culture at the level of country, based on where participants were raised. It is an open question whether performance in our participant populations (of relatively young and well-educated adults) is representative of the broader country. This is especially true for societies with substantial ethnic and cultural variation such as the US. We expect that our data is likely to underestimate variation both within and between the countries we sample from.

@@

<!-- Additionally, outside of language, cross-cultural differences could come from a couple different mechanisms:  -->
In this study, we ask whether variation in lexical co-occurrence can account for the cross-context (culture and language) differences observed in how people evaluate similarity. First, we attempt to replicate @Ji2004 by running a study comparing English speakers in the US and Mandarin speakers in mainland China. Second, we evaluate whether the differences observed between English and Mandarin speakers’ similarity judgments also extend to comparisons between English and Vietnamese, by running a similar study between English speakers in the US and Vietnamese speakers in Vietnam. Vietnam is a Southeast Asian country that borders China and is historically greatly influenced by Chinese culture [@Hui2002]. Therefore, it serves as a suitable cultural context to investigate whether the claim made by @Ji2004 and previous studies, that Eastern and Western cultures have different notions of similarity, extends beyond mainland China.

@@

# Methods 
```{r get responding and demographic data, include = F}
df <- read.csv("../../../data/data_USCNVN_ENZHVI.csv")
#df <- read.csv("../../../data/data_USCNVN_ENZHVI_0623.csv")

@@

triads_omit <- read.csv("../../../data/triads_omit.csv") %>%
  pull(x)

head(df)
```

@@

We excluded `r df.excl_US$ppts_finished - df.excl_US$ppts_after_att_check_excl` US participants, `r df.excl_VN$ppts_finished - df.excl_VN$ppts_after_att_check_excl` Vietnam participants, and `r df.excl_CN$ppts_finished - df.excl_CN$ppts_after_att_check_excl` China participants who did not answer all attention checks correctly. We followed 4 exclusion criteria that aim to retain only participants who are influenced by one culture: (1) non-native speakers of English and Vietnamese, respectively, (2) fluent in at least one of the other two study languages (Vietnamese for US participants, English for Vietnamese participants and Chinese participants), (3) have lived outside of the test country (US, Vietnam, or China) for more than two years, and (4) have significant international experience (more than 6 international experiences of 2 days or longer.) We did not use a particular criterion for a language if it would exclude 25% or more of any one sample. In this round of exclusion, we excluded `r df.excl_US$ppts_after_att_check_excl - df.excl_US$ppts_after_demog_excl` US, `r df.excl_VN$ppts_after_att_check_excl - df.excl_VN$ppts_after_demog_excl` Vietnam participants, and `r df.excl_CN$ppts_after_att_check_excl - df.excl_CN$ppts_after_demog_excl` China participants. After these exclusions, the US sample included `r df.excl_US$ppts_after_demog_excl` participants (`r df.gender_E %>% filter(gender == "Male") %>% pull(n)`M, `r df.gender_E %>% filter(gender == "Female") %>% pull(n)`F, `r df.gender_E %>% filter(gender == "Non-binary") %>% pull(n)` non-binary, `r df.gender_E %>% filter(gender == "Decline to answer") %>% pull(n)` other), with mean age = `r round(df.age_E$mean_age, 2)` (SD = `r round(df.age_E$sd_age, 2)`) and median age = `r df.age_E$median_age`. The Vietnam sample included `r df.excl_VN$ppts_after_demog_excl` participants (`r df.gender_V %>% filter(gender == "Nam") %>% pull(n)`M, `r df.gender_V %>% filter(gender == "Nữ") %>% pull(n)`F, `r df.gender_V %>% filter(gender == "Từ chối trả lời") %>% pull(n)` other), with mean age = `r round(df.age_V$mean_age, 2)` (SD = `r round(df.age_V$sd_age, 2)`) and median age = `r df.age_V$median_age`. The China sample included `r df.excl_CN$ppts_after_demog_excl` participants (`r df.gender_M %>% filter(gender == "男性") %>% pull(n)`M, `r df.gender_M %>% filter(gender == "女性") %>% pull(n)`F, `r df.gender_M %>% filter(gender == "拒绝回答") %>% pull(n)` other), with mean age = `r round(df.age_M$mean_age, 2)` (SD = `r round(df.age_M$sd_age, 2)`) and median age = `r df.age_M$median_age`.[^footnote]

\footnote \smallfont A table summarizing number of participants lost at each round of exclusions is included in the Appendix.

Notably, we lost a majority of our Vietnam participants (more than 60%) in the attention check exclusion. One reason we suspect why this might have happened is because our Vietnam participants are less familiar with research surveys and attention check questions, and thus might have thought too much about the attention check questions.[^footnote]

\footnote \smallfont We carried out an exploratory analysis where we used a less stringent attention check exclusion (covered in the final part of the Results & Analysis section). To preview our findings for this analysis, we did not find any difference in significant results when compared with the main analysis.

@@

Mandarin co-occurrence was collected using the online interface of the Chinese Lexical Association Database (Lin et al, 2019). Mandarin also makes frequent use of compositional morphology. To match the Vietnamese corpus statistics, we included in our counts any instances of the target term or close semantic neighbors containing the same morpheme(s) following the rule described above. These semantic neighbors were either suggested to us by the online interface (which gave search suggestions for words beginning with the target term as we typed them in) or suggested by a native speaker of Mandarin conducting the co-occurrence count data collection. For example, in getting the collocates of the term "門" (door), we also included "門兒" (also door, but is returned as a different word). Compounds that do not meet this criteria were excluded from collocate counts. For example "門牙" (door.tooth) refers to the front tooth even though it includes the same character as the target word "門" (door), so was therefore excluded from our collocate counts for "door".

@@

fit.country_EN_VN = glmer(responses_theme ~ country + (1 | subject) + (country | triad), 
                    data = df %>% filter(country != "China"),
                    family = "binomial")
summary(fit.country_EN_VN) 
```
The group means of proportion of thematic response in mainland China is the highest (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by the groups means in Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), which is slightly higher than that of the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`) (Fig 1).

```{r echo=FALSE, warning=FALSE, cache = TRUE, fig.cap="Proportion of thematic responses by country."}

#show violin plot
ggplot(df.country,
       mapping = aes(x = factor(country, levels=c("China", "US", "Vietnam")), 
                     y = theme_resp_percent, 
                     color = factor(country, levels=c("China", "US", "Vietnam")))) +
  geom_violin() +
  geom_jitter(height = 0, 
              alpha = 0.3) +  
  stat_summary(fun.data = "mean_cl_boot", 
               geom = "pointrange") + 
  labs(y = "Proportion Thematic Chosen", 
       x = "Country") + 
  scale_color_manual(values=c("#D63230", "#1C77C3", "#F39237")) + 
  theme(legend.position = "none")
```

@@

## 2. Can the differences in similarity judgments between English, Mandarin and Vietnamese speakers be explained by variation in lexical statistics?
### Is responding in each cultural context predicted by the corresponding lexical statistics (raw co-occurrences)? 
#### Raw lexical co-occurrences

@@

```{r US responding predicted by each corpus model raw co-occurrences, warning=FALSE, include=F, cache = TRUE}
# English 
fit.EN_US = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)

@@

#Mandarin
fit.ZH_US = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_US)
fit.ZH_US.anova = Anova(fit.ZH_US, type = 3)
```

@@

#Mandarin
fit.ZH_VN = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_VN)
fit.ZH_VN.anova = Anova(fit.ZH_VN, type = 3)
```

@@

```{r calculate range of beta statistics, warning=F, include=F}
beta.single_corpus_freq <- c(
  round(fixef(fit.EN_US)["theme_freq_prop_E"], 2),
  round(fixef(fit.VI_US)["theme_freq_prop_V"], 2), 
  round(fixef(fit.ZH_US)["theme_freq_prop_M"], 2),

  round(fixef(fit.EN_CN)["theme_freq_prop_E"], 2),
  round(fixef(fit.VI_CN)["theme_freq_prop_V"], 2), 
  round(fixef(fit.ZH_CN)["theme_freq_prop_M"], 2),

  round(fixef(fit.EN_VN)["theme_freq_prop_E"], 2),
  round(fixef(fit.VI_VN)["theme_freq_prop_V"], 2), 
  round(fixef(fit.ZH_VN)["theme_freq_prop_M"], 2)

)
```
We found that all corpora are significant predictors of all cultural context responding, with p < 0.05 and $\beta$ from `r min(beta.single_corpus_freq)` - `r max(beta.single_corpus_freq)`. (For a full report, see Appendix.)

@@

@@

#Mandarin
fit.ZH_US_cos = glmer(responses_theme ~ theme_cosine_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US"), 
          family="binomial")
summary(fit.ZH_US_cos)
fit.ZH_US_cos.anova = Anova(fit.ZH_US_cos, type = 3)
fit.ZH_US_cos.anova
```

@@

@@

#### Raw co-occurrences

@@

We observed some level of language specificity from this analysis. The English corpus is the only significant predictor for US responding, and the Mandarin corpus is the only significant predictor for China response. However, against our prediction, the Vietnamese corpus was not a significant predictor for Vietnam responding when other corpora were included. (Figure 2a).
